---
title: "project proposal"
author: "Mehdi Khan"
date: "October 25, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=65),tidy=TRUE)
```

load libraries:
```{r}
suppressMessages(suppressWarnings(library(dplyr))) 
suppressMessages(suppressWarnings(library(stringr))) 
suppressMessages(suppressWarnings(library(DATA606))) 
suppressMessages(suppressWarnings(library(psych)))
suppressMessages(suppressWarnings(library(ggplot2)))
```

#data prepration: 

Load data:
```{r}
originDS <- read.csv("https://raw.githubusercontent.com/kmehdi2017/projectProp/master/ProjectProposal/ACS_15_5YR_DP02_with_ann.csv",sep = ",",stringsAsFactors = FALSE)

head(originDS)


```

The original data contains a lot of variable which are not relevant to this study, therefore a subset of the data is extracted. The following are the variables that are selected for the project as described below:

GEO.display.label: Geography
HC01_VC03	Estimate; HOUSEHOLDS BY TYPE - Total households
HC01_VC04:	Estimate: Total households - Family households (families)	
HC01_VC06:	Estimate: Total households - Family households (families) - Married-couple family										
HC01_VC08:	Estimate: Total households - Family households (families) - Male householder, no wife present	
HC01_VC10:	Estimate: Total households - Family households (families) - Female householder, no husband present 
HC01_VC76:	Estimate: SCHOOL ENROLLMENT - Population 3 years and over enrolled in school
HC01_VC88:	Estimate: EDUCATIONAL ATTAINMENT - Population 25 years and over - High school graduate
HC03_VC88:	Percent:  EDUCATIONAL ATTAINMENT - Population 25 years and over - High school graduate 
HC01_VC91:	Estimate: EDUCATIONAL ATTAINMENT - Population 25 years and over - Bachelor's degree
HC03_VC91:	Percent; EDUCATIONAL ATTAINMENT - Population 25 years and over - Bachelor's degree

 								
```{r}


vars <- c("GEO.display.label","HC01_VC03","HC01_VC04","HC01_VC06" , "HC01_VC08",  "HC01_VC10",   "HC01_VC76", "HC01_VC88","HC03_VC88", "HC01_VC91","HC03_VC91")

familyEduDS <- originDS[-1,vars]

head(familyEduDS)

```

providing meaningful names to columns:
```{r}

columnNames <- c("states","total_house_hold","total_family","married_couple_family" , "husband_only_family",  "wife_only_family",   "school_enrollment", "high_school_grad", "percent_high_school", "bachelor_deg_holder","percent_bachelor")

colnames(familyEduDS) <- columnNames

head(familyEduDS)

```

converting  column type to numeric for columns holding numbers:
```{r}

familyEduDS[2:11] <- sapply(familyEduDS[2:11],as.numeric)

```


Calculate percentage of types of families in relation to total numaber of families:
```{r}
percent_married_couple <- familyEduDS$married_couple_family/familyEduDS$total_house_hold
percent_husband_only <- familyEduDS$husband_only_family/familyEduDS$total_house_hold
percent_wife_only <- familyEduDS$wife_only_family/familyEduDS$total_house_hold

```

## Research question
Does larger number of married couple families cause larger number of educated population compared to the number of families of single parents?

## case
Each case represents a state in the United States, there are 51 of them. 

## Data collection
The  data represents selected social chracterstics in the United States in the period of 5 years from  2011 to 2015 and compiled by American Community Survey,US Census Bureau, the premier source for detailed information about the American people and workforce.

The data was published by US Census Bureau and posted in American Fact Finder website: https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_15_5YR_DP02&src=pt
 


## Type of study
This is an observational study

## Data Source
The data was published by US Census Bureau and posted in American Fact Finder website: https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_15_5YR_DP02&src=pt

##Response
he response variables are the estimates of school enrollment, high school gradutes and bachelor degree and they are numerical.

##Explanatory
The explanatory variables are the estimates of family types and are numerical.


## Relevant summary statistics
```{r}
describe(percent_married_couple)

```

```{r}
describe(percent_husband_only)

```

```{r}
describe(percent_wife_only)

```

```{r}
describe(familyEduDS$school_enrollment)

```

```{r}
describe(familyEduDS$high_school_grad)

```


```{r}
describe(familyEduDS$bachelor_deg_holder)

```

```{r}
describe(familyEduDS$percent_high_school)

```

```{r}
describe(familyEduDS$percent_bachelor)

```

```{r}
qplot(percent_married_couple, geom="histogram") 
```





