---
title: "project proposal"
author: "Mehdi Khan"
date: "October 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries:
```{r}
suppressMessages(suppressWarnings(library(dplyr))) 
suppressMessages(suppressWarnings(library(stringr))) 
suppressMessages(suppressWarnings(library(DATA606))) 
suppressMessages(suppressWarnings(library(psych)))
suppressMessages(suppressWarnings(library(ggplot2)))
```

#data: 
data source: American Fact Finder, US Census Bureau, https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_15_5YR_DP02&src=pt

Load data:
```{r}
originDS <- read.csv("https://raw.githubusercontent.com/kmehdi2017/projectProp/master/ProjectProposal/ACS_15_5YR_DP02_with_ann.csv",sep = ",",stringsAsFactors = FALSE)
```

The original data contains a lot of variable which are not relevant to this project, therefore a subset of the data is extracted. The following are the variables that are selected for the project and described below:

HC01_VC03:	Estimate; Total households	
HC01_VC04:	Estimate; Total households - Family households (families)	
HC01_VC06:	Estimate; Total households - Family households (families) - Married-couple family										
HC01_VC08:	Estimate; Total households - Family households (families) - Male householder, no wife present	
HC01_VC10:	Estimate; Total households - Family households (families) - Female householder, no husband present 
HC01_VC12:	Estimate; Total households - Nonfamily households		
HC01_VC13:	Estimate; Total households - Nonfamily households - Householder living alone	
HC01_VC21:	Estimate;  Average household size
HC01_VC22:	Estimate;  Average family size	
HC01_VC76:	Estimate; SCHOOL ENROLLMENT - Population 3 years and over enrolled in school	
HC01_VC85:	Estimate; EDUCATIONAL ATTAINMENT - Population 25 years and over	
HC01_VC88:	Estimate; EDUCATIONAL ATTAINMENT - Population 25 years and over - High school graduate
HC01_VC89:	Estimate; EDUCATIONAL ATTAINMENT - Population 25 years and over - Some college, no degree
HC01_VC90:	Estimate; EDUCATIONAL ATTAINMENT - Population 25 years and over - Associate's degree
HC01_VC91:	Estimate; EDUCATIONAL ATTAINMENT - Population 25 years and over - Bachelor's degree
HC01_VC92:	Estimate; EDUCATIONAL ATTAINMENT - Population 25 years and over - Graduate or professional degree
HC01_VC95:	Estimate; EDUCATIONAL ATTAINMENT - Percent high school graduate or higher	
HC01_VC96:  Estimate; EDUCATIONAL ATTAINMENT - Percent bachelor's degree or higher
HC01_VC136:	Estimate; PLACE OF BIRTH - Total population - Foreign born
HC01_VC140:	Estimate; U.S. CITIZENSHIP STATUS - Foreign-born population	
HC01_VC141:	Estimate; U.S. CITIZENSHIP STATUS - Foreign-born population - Naturalized U.S. citizen								
```{r}
head(originDS)

vars <- c("GEO.display.label","HC01_VC03","HC01_VC04","HC01_VC06" , "HC01_VC08",  "HC01_VC10",  "HC01_VC12",  "HC01_VC13", "HC01_VC21",  "HC01_VC22",  "HC01_VC76",  "HC01_VC85",  "HC01_VC88",  "HC01_VC89",  "HC01_VC90", 
"HC01_VC91",  "HC01_VC92",  "HC01_VC95",  "HC01_VC96",  "HC01_VC136", "HC01_VC140", "HC01_VC141")

familyEduDS <- originDS[-1,vars]


```

providing meaningful names to columns:
```{r}

columnNames <- c("states","tot_hh","fam_hh","married_couple" , "husband_only",  "wife_only",  "non_fam_hh",  "non_fam_alone", "avg_hh_size",  "avg_fam_size",  "sch_enrollment",  "edu_attainment",  "high_sch_grad",  "some college",  "associate_deg", "bachelor_deg",  "graduate_prof",  "percent_highsch",  "percent_bachelor",  "tot_foreign_born", "citizen_foreign", "natur_citizen")

colnames(familyEduDS) <- columnNames

```

converting  column type to numeric for columns holding numbers:
```{r}


familyEduDS[2:22] <- sapply(familyEduDS[2:22],as.numeric)

```

calucalating average values of some variables:
```{r}
mean_tot_hh <- mean(familyEduDS$tot_hh)
mean_fam_hh <- mean (familyEduDS$fam_hh)
mean_couple <- mean(familyEduDS$married_couple)
mean_husband_only <- mean(familyEduDS$husband_only)
mean_wife_only <- mean(familyEduDS$wife_only)
mean_high_school <- mean(familyEduDS$high_sch_grad)
mean_bachelor <-  mean(familyEduDS$bachelor_deg)

```

ratio of education and family types
```{r}
percent_HS_couple <- familyEduDS$high_sch_grad/familyEduDS$married_couple
percent_HS_husband_Only <- familyEduDS$high_sch_grad/familyEduDS$husband_only
percent_HS_wife_Only <- familyEduDS$high_sch_grad/familyEduDS$wife_only

percent_BA_couple <- familyEduDS$bachelor_deg/familyEduDS$married_couple
percent_BA_husband_Only <- familyEduDS$bachelor_deg/familyEduDS$husband_only
percent_BA_wife_Only <- familyEduDS$bachelor_deg/familyEduDS$wife_only

percent_enroll_couple <- familyEduDS$sch_enrollment/familyEduDS$married_couple
percent_enroll_husband_Only <- familyEduDS$sch_enrollment/familyEduDS$husband_only
percent_enroll_wife_Only <- familyEduDS$sch_enrollment/familyEduDS$wife_only

```


some statistics:
```{r}
describe(familyEduDS$married_couple)

describe(familyEduDS$bachelor_deg)

```

Histograms:
```{r}
ggplot(data.frame(percent_HS_couple), aes(x=percent_HS_couple, count(percent_HS_couple)))+geom_line()
stat_bin(60)
qplot(percent_HS_couple, geom="histogram") 
```





```{r}

qplot(percent_HS_husband_Only, geom="histogram") 

```



